VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-19 snapshot_system/paper_background.tex
change del

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter to identify non-duplicates and a prefetching cache for potential
duplicates that might hit in the future.
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in ~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
\comments{
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
}
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

@@

\comments{
For example, the
sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011},
uses 5GB memory space for 100TB of raw data even the compressed data size
is about 5TB. It is fairly typical for
a physical machine with terabytes of disk storage and tens of gigabytes of memory
to support  more than 20 VMs and 10 snapshots per VM and can require  over 100TB of raw space
even the actual storage space need is about 5TB assuming 20:1 compression ratio with deduplication.
It is not easy  to adopt the sampled index method  for a distributed architecture
and every deduplication request may access a remote machine for index lookup and the overall overhead of access latency 
for all requests can be significant. 
 In general
with a cluster of machines in a cloud cluster, there will be more memory requirement per machine such as a 
fingerprint cache, for cross-machine fingerprint comparison~\cite{bottleneck08,Dong2011} and a deduplication
task often needs several gigabytpes of memory, which impacts other cloud services sharing the computing resource.
}

@@

\comments{
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

Indeed, the greater the deduplication efficiency, the less
the data redundancy is maintained, and the greater the ``hidden'' data
dependence between VMs that must be hosted as if they were physically isolated.
Thus,
restricting deduplication to be local to each VM (but across snapshots) 
aligns the units of storage with the isolation properties guaranteed by the
cloud at the possible expense of storage efficiency.
%Localizing the impact of deduplication can increase fault isolation and resilience.
%Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
%discouraged.

Additionally, storage efficient deduplication can be computationally expensive
to implement in a distributed setting.  In particular, duplicate detection is
logically a ``global'' operation over all data chunks
under consideration.
By localizing duplicate search, the computational requirements
for deduplication can be reduced, because fewer chunks must be compared.
Further, deduplication of VMs can easily be implemented in
parallel.

Another disadvantage of cross-VM sharing is that it complicates 
snapshot deletion, which occurs frequently because snapshots expire regularly.
Several well developed garbage collection methods do not apply to the
converged virtualized architecture due to various environment and resource limitations:
Grouped mark-and-sweep~\cite{Guo2011} is an effective approach by tracking
the dependencies from data containers to metadata groups, such that only a subset of
metadata needs to be scanned in order to reclaim the garbage in any given data container.
However, data in VM environment tend to be highly dedupable which means
such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
 especially in a distributed setting.
Perfect hashing~\cite{Fabiano2013} provides a very memory efficient way to represent a large set of
hashes in a compact form, however, it still requires the scan of all data hashes before any garbage could begin,
which limits its application to only offline batch garbage collection. In addition, it requires a centralized location
to store the hashing result which is still too big to be hold anywhere in a converged architecture.

%to count if a data chunk is still shared by other snapshots.

Thus,
localizing deduplication to the VM can align fault isolation properties and
guarantees, improve the computational efficiency of snapshot maintenance, and
simplify snapshot deletion over non-localized approaches at the possible
expense of storage efficiency.

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block  size in a distributed file system such as  
the Hadoop file system and the Google file system~\cite{googlefs03}
is large (e.g.  64MB) so as to optimize file system performance
for big data.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average~\cite{Jin2009}).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for snapshot storage.

%a minimum association of FSBs to VMs in the packaging
%process. By minimizing this association we can improve fault tolerance by
%reducing the number of VMs affected when storage nodes are lost.

\paragraph*{Moderating Resource Usage} --
In a converged setting, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service must share the resources available to
other cloud services as illustrated in Figure~\ref{fig:collocated},
and has to minimize its impact on user-deliverable performance.
%The key resource for global comparison  is memory for storing the fingerprints.
%During snapshot deletion, reference counting for all unique blocks used in all VMs requires a significant resource usage
%too.  Thus we seek for the low-profile techniques with approximation and low resource consumption.

}

\comments{
With the above  considerations in mind, we propose a 
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.  
This compares the traditional deduplication approach {\em VM-oblivious} (VO)
which manages duplicate data chunks without consideration of VMs.
We discuss our key ideas and design in the next section.
%Our key idea is to simplify deduplication management and 
}
\comments{
In designing a VC deduplication solution, we have considered and adopted some of
the following previously-developed techniques.
1) {\em Changed Block Tracking}. 
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized to the snapshot
history for each VM.
%We are seeking for a tradeoff since 
%global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
1) {\em Stateless  Data Routing}.
One approach to scalable duplicate comparison is to use content-based hash
partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index for Memory Parsimony}. 
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation 
and discuss snapshot deletion support, then present an implementation design. 
}

>>>>

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter to identify non-duplicates and a prefetching cache for potential
duplicates that might hit in the future.
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in ~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
% \comments{
% All of the above approaches have focused on optimization of deduplication
% efficiency or speed, and none of them have considered the impact
% of deduplication on fault resilience in a cloud environment.  To do so, we
% examine the following properties of VM snapshot deduplication for converged
% IaaS architectures.
% }
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

@@

% \comments{
% For example, the
% sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011},
% uses 5GB memory space for 100TB of raw data even the compressed data size
% is about 5TB. It is fairly typical for
% a physical machine with terabytes of disk storage and tens of gigabytes of memory
% to support  more than 20 VMs and 10 snapshots per VM and can require  over 100TB of raw space
% even the actual storage space need is about 5TB assuming 20:1 compression ratio with deduplication.
% It is not easy  to adopt the sampled index method  for a distributed architecture
% and every deduplication request may access a remote machine for index lookup and the overall overhead of access latency
% for all requests can be significant.
%  In general
% with a cluster of machines in a cloud cluster, there will be more memory requirement per machine such as a
% fingerprint cache, for cross-machine fingerprint comparison~\cite{bottleneck08,Dong2011} and a deduplication
% task often needs several gigabytpes of memory, which impacts other cloud services sharing the computing resource.
% }

@@

% \comments{
% \paragraph*{Local versus Global Deduplication} --
% If deduplication is implemented strictly in the storage system,
% a data chunk is compared with fingerprints collected from all VMs during
% the deduplication process, and only a small number of
% copies of duplicates are stored in the storage.  This sharing
% artificially creates hidden data dependencies among different VMs owned by
% different
% who
% contract for isolation
% by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
% user is the VM, but the unit of sharing in a deduplicated data store is the
% data chunk.

% Indeed, the greater the deduplication efficiency, the less
% the data redundancy is maintained, and the greater the ``hidden'' data
% dependence between VMs that must be hosted as if they were physically isolated.
% Thus,
% restricting deduplication to be local to each VM (but across snapshots)
% aligns the units of storage with the isolation properties guaranteed by the
% cloud at the possible expense of storage efficiency.
% %Localizing the impact of deduplication can increase fault isolation and resilience.
% %Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is
% %discouraged.

% Additionally, storage efficient deduplication can be computationally expensive
% to implement in a distributed setting.  In particular, duplicate detection is
% logically a ``global'' operation over all data chunks
% under consideration.
% By localizing duplicate search, the computational requirements
% for deduplication can be reduced, because fewer chunks must be compared.
% Further, deduplication of VMs can easily be implemented in
% parallel.

% Another disadvantage of cross-VM sharing is that it complicates
% snapshot deletion, which occurs frequently because snapshots expire regularly.
% Several well developed garbage collection methods do not apply to the
% converged virtualized architecture due to various environment and resource limitations:
% Grouped mark-and-sweep~\cite{Guo2011} is an effective approach by tracking
% the dependencies from data containers to metadata groups, such that only a subset of
% metadata needs to be scanned in order to reclaim the garbage in any given data container.
% However, data in VM environment tend to be highly dedupable which means
% such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
%  especially in a distributed setting.
% Perfect hashing~\cite{Fabiano2013} provides a very memory efficient way to represent a large set of
% hashes in a compact form, however, it still requires the scan of all data hashes before any garbage could begin,
% which limits its application to only offline batch garbage collection. In addition, it requires a centralized location
% to store the hashing result which is still too big to be hold anywhere in a converged architecture.

% %to count if a data chunk is still shared by other snapshots.

% Thus,
% localizing deduplication to the VM can align fault isolation properties and
% guarantees, improve the computational efficiency of snapshot maintenance, and
% simplify snapshot deletion over non-localized approaches at the possible
% expense of storage efficiency.

% %minimize data sharing and simplify deletion while sacrificing
% %deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
% \paragraph*{Units of Storage versus Units of Sharing} --
% The file system block  size in a distributed file system such as
% the Hadoop file system and the Google file system~\cite{googlefs03}
% is large (e.g.  64MB) so as to optimize file system performance
% for big data.
% Typically, deduplication implementations use smaller block sizes (4K bytes to
% 8K bytes on average~\cite{Jin2009}).
% In addition to the failure correlation that cross-VM data deduplication can
% introduce, disk block sharing of data chunks is another potential source of
% hidden data dependence.
% Thus a VM-centric solution should consider how data chunks from VMs are
% distributed among large file system blocks when a large-scale distributed file
% system is used for snapshot storage.

% %a minimum association of FSBs to VMs in the packaging
% %process. By minimizing this association we can improve fault tolerance by
% %reducing the number of VMs affected when storage nodes are lost.

% \paragraph*{Moderating Resource Usage} --
% In a converged setting, the resources that are used to implement snapshot
% backup and deduplication are the same resources that must support cloud-hosted
% VMs.  Thus the backup service must share the resources available to
% other cloud services as illustrated in Figure~\ref{fig:collocated},
% and has to minimize its impact on user-deliverable performance.
% %The key resource for global comparison  is memory for storing the fingerprints.
% %During snapshot deletion, reference counting for all unique blocks used in all VMs requires a significant resource usage
% %too.  Thus we seek for the low-profile techniques with approximation and low resource consumption.

% }

% \comments{
% With the above  considerations in mind, we propose a
% VM-centric approach (called VC)
% for a co-located backup service that has a resource usage profile
% suitable for use with converged cloud architectures.
% This compares the traditional deduplication approach {\em VM-oblivious} (VO)
% which manages duplicate data chunks without consideration of VMs.
% We discuss our key ideas and design in the next section.
% %Our key idea is to simplify deduplication management and
% }
% \comments{
% In designing a VC deduplication solution, we have considered and adopted some of
% the following previously-developed techniques.
% 1) {\em Changed Block Tracking}.
% VM snapshots can be  backed up  incrementally by identifying data segments that have
% changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
% Such a scheme  is  VM-centric since deduplication is localized to the snapshot
% history for each VM.
% %We are seeking for a tradeoff since
% %global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
% 1) {\em Stateless  Data Routing}.
% One approach to scalable duplicate comparison is to use content-based hash
% partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
% Stateless data routing divides the deduplication work with a similarity approximation. This work
% is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and
% each request is routed  to a machine which holds
% a Bloom filter  or can fetch on-disk index for additional comparison.
% While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
% VM-specific local search.
% 3) {\em Sampled Index for Memory Parsimony}.
% One effective approach that reduces memory usage is
% to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}.
% The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture.
% To use a distributed memory version of the sampled index, every deduplication request
% may access a remote machine for index lookup and the overall overhead of access latency for all requests
% can be significant.

% We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation
% and discuss snapshot deletion support, then present an implementation design.
% }

### 2014-07-19 snapshot_system/paper_del.tex
re
change del

Snapshot deletions can occur frequently since old snapshots become less useful.
General deduplication complicates the deletion process because sharing of duplicates
requires a global reference counting~\cite{Guo2011,Fabiano2013} 
 to identify if  a chunk can be  safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, since
sharing in our approach is limited to only a small dataset, we can greatly simplify deletion by focusing
on intra-VM sharing rather than inter-VM sharing. Then the snapshot
 deletion can therefore be conducted independently among VMs.
\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.

>>>>

Snapshot deletions can occur frequently since old snapshots become less useful.
Both space accounting and reclaimation are difficult because deduplication
greatly complicates how we identify unused data when deletion occurs.
To filter out un-referenced data from shared duplicates,
it would require to either maintain expensive live schemas and operations
such as global reference counting,
or use some advanced mark-and-sweep methodologies~\cite{Guo2011,Fabiano2013},
which are all unsuitable for converged architecture as discussed in section~\ref{sect:background}.
% to identify if a chunk can be safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, since
sharing in our approach is limited to only a small dataset, we can greatly simplify deletion by focusing
on intra-VM sharing rather than inter-VM sharing. Then the snapshot
 deletion can therefore be conducted independently among VMs.
\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.
